---
title: "EEPC Aggregate Loss Model:DRAFT"
author: "CDEEO"
date: "Thursday, July 02, 2015"
output: html_document
---
```{r,echo=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE,message=F)
```

###Data Source: [EEPC Annual Reports 2002-2010][EEPC]
###Summary Statistics on Cost
1. Overall   
```{r}
EEPC <- read.csv("../Data/EEPC Risk Analysis.csv", stringsAsFactors=FALSE)
Total <- read.csv("../Data/EEPC Total.csv", stringsAsFactors=FALSE)
set.seed(123)

library(moments)
library(MASS)
library(dplyr)
library(GLDEX)

EEPC %>% summarise(mean=mean(Amount), standard_dev=sd(Amount), 
                   median= median(Amount),skewness=skewness(Amount),
                   kurtosis=kurtosis(Amount))

```

2. By Year
```{r,fig.height=12,fig.width=12}
EEPC %>% group_by(year.settled) %>% 
    summarise(mean=mean(Amount), total=sum(Amount), standard_dev=sd(Amount),
              median= median(Amount),skewness=skewness(Amount),
              kurtosis=kurtosis(Amount))
par(mfrow=c(3,3))
for(y in 2002:2010){
    cost=EEPC[EEPC$year.settled==y,"Amount"]
    hist(cost,col=sample(colors(),1),breaks = 25,
         main=y,xaxt="n",
         xlab="EEPC Law Suit Cost (in thousands)",
         xlim=c(0,quantile(cost,.99)))
    xrange=seq.int(0,quantile(cost,.99),length.out = 6)
    xlabel=paste("$",round(xrange/1e3),sep="")
    axis(1,at=xrange,labels = xlabel)
}

```

             
###Severity Loss Probability Model   
**1. Exponential Distribution**       
```{r}
N=nrow(EEPC)
exp_rate=fitdistr(EEPC$Amount,densfun = "exponential")
exp_rate$estimate
par(mfrow=c(1,1))
hist(EEPC$Amount,xlim = c(0,1e6),breaks = 5e2,freq = F,
     main="Histogram with Exponential Distribution",
     col="green",xaxt="n",yaxt="n",
     xlab="EEPC Law Suit Cost (in thousands)")
xrange=seq.int(0,1e6,by = 100)
yexp=dexp(xrange,exp_rate$estimate)
lines(x = xrange,y = yexp,lwd=3,col="blue")
xrange=seq.int(0,1e6,by = 2e5)
xlabel=paste("$",xrange/1000,sep="")
axis(1,at=xrange,labels = xlabel)
yrange=seq.int(0,1.2e-5,by = 2e-6)
ylabel=paste(yrange*100,"%",sep="")
axis(2,at=yrange,labels = ylabel)


```

###Diagnostic against Exponential Model  
**QQplot**    
```{r}
rand_exp=rexp(1e5,rate = exp_rate$estimate)
exprange=c(0,1e6)
qqplot(EEPC$Amount,rand_exp,col="darkgrey",pch=19,xlim=exprange,ylim=exprange,
       ylab="Exponential Distribution",xlab="EEPC Law Suit Costs",main="QQ-plot")
abline(0,1,col="red",lty=3,lwd=2)
```

###Kolmogorov-Smirnov (K-S) test   
```{r}
ks.gof(EEPC$Amount,"pexp",rate = exp_rate$estimate)
```

**Both diagnostics show that there is a signifcant departure from the Exponential Model** 


###2. LogNormal Distribution
```{r}
EEPC$LogAmount=log(EEPC$Amount)
lnorm_rate=fitdistr(EEPC$Amount,densfun = "log-normal")
lnorm_rate$estimate
hist(EEPC$Amount,xlim = c(0,1e6),breaks = 5e2,freq = F,
     main="Histogram with Log-Normal Distribution",
     col="yellow",xaxt="n",yaxt="n",ylim=c(0,1.4e-5),
     xlab="EEPC Law Suit Cost (in thousands)")
xrange=seq.int(0,1e6,by = 100)
ylog=dlnorm(xrange,meanlog = lnorm_rate$estimate["meanlog"],
            sdlog = lnorm_rate$estimate["sdlog"])
lines(x = xrange,y = ylog,lwd=3,col="purple")
xrange=seq.int(0,1e6,by = 2e5)
xlabel=paste("$",xrange/1000,sep="")
axis(1,at=xrange,labels = xlabel)
yrange=seq.int(0,1.4e-5,by = 2e-6)
ylabel=paste(yrange*100,"%",sep="")
axis(2,at=yrange,labels = ylabel)


norm_rate=fitdistr(EEPC$LogAmount,densfun = "normal")
hist(EEPC$LogAmount,freq = F,xlim = c(6,14),breaks = 20,
     main="Histogram with Normal Distribution",
     col="pink",yaxt="n",
     xlab="EEPC Law Suit Cost (Log Scale)")
xrange=seq.int(0,18,by = .1)
ynorm=dnorm(xrange,mean = norm_rate$estimate["mean"],
            sd = norm_rate$estimate["sd"])
lines(x = xrange,y = ynorm,lwd=3,col="darkgreen")

yrange=seq.int(0,0.4,by = 0.05)
ylabel=paste(yrange*100,"%",sep="")
axis(2,at=yrange,labels = ylabel,las=2,cex=.5)

```

###Diagnostic   

```{r}
rand_lnorm=rlnorm(1e5,meanlog = lnorm_rate$estimate["meanlog"],
                 sdlog = lnorm_rate$estimate["sdlog"])
exprange=c(0,1e6)
qqplot(EEPC$Amount,rand_lnorm,
       col="darkblue",pch=19,xlim=exprange,ylim=exprange,
       ylab="LogNormal Distribution",xlab="EEPC Law Suit Costs",main="QQ-plot")
abline(0,1,col="red",lty=3,lwd=3)
```

###Kolmogorov-Smirnov (K-S) test   
```{r}
ks.gof(EEPC$Amount,"plnorm",meanlog = lnorm_rate$estimate["meanlog"],
       sdlog = lnorm_rate$estimate["sdlog"])
```

**Diagnostics show that there is not a significant level of evidence to reject the LogNormal Model for Severity of Loss.**       

###Loss Frequence Model
             
```{r}
year=EEPC %>% group_by(year.settled) %>% summarise(AvgAmount=mean(Amount),N=n())
year %>% summarise(Avg=mean(N),Var=var(N),Median=median(N),
                   skewness(N),kurtosis(N))
barplot(year$N,names.arg =year$year.settled,col=sample(colors(),1),
        xlab="year",main="Number of cases settled")
```

**Because sample variance is much larger than sample mean: use Negative Binomial**                
```{r}
nb_param=fitdistr(year$N,"Negative Binomial")
nb_param$estimate
```


###Kolmogorov-Smirnov (K-S) test   
```{r}
ks.gof(year$N,"pnbinom",size = nb_param$estimate["size"],
       mu = nb_param$estimate["mu"])
```

**Diagnostics show that there is not a significant level of evidence to reject the Negative Binomial Model for Loss Frequency.**     

###Aggregated Loss Model using Monte Carlo Simulation

```{r}
total_cost=rep(0,5e4)
for(j in 1:5e4){
    num_cases=rnbinom(n=1,size = nb_param$estimate["size"],
                      mu = nb_param$estimate["mu"])
    law_suit_cost=0
    for(i in 1:num_cases){
        law_suit_cost=law_suit_cost+
            rlnorm(n = 1,meanlog=lnorm_rate$estimate["meanlog"],
                   sdlog = lnorm_rate$estimate["sdlog"])
    }
    total_cost[j]=law_suit_cost
}

summary(total_cost)
lnorm_param=fitdistr(total_cost,densfun = "log-normal")
lnorm_param$estimate

hist(total_cost,breaks = 50,freq = F,
     main="Aggregated Loss Distribution",
     col="grey",ylim=c(0,2.5e-7),xaxt="n",#yaxt="n",
     xlab="Total Cost (in millions)")
xrange=seq.int(0,1e8,by = 100)
ylnorm=dlnorm(xrange,meanlog = lnorm_param$estimate["meanlog"],
            sdlog = lnorm_param$estimate["sdlog"])
lines(x = xrange,y = ylnorm,lwd=3,col="darkgreen")
xrange=seq.int(0,3.5e7,by = 5e6)
xlabel=paste("$",xrange/1e6,sep="")
axis(1,at=xrange,labels = xlabel)

```        

###Comparsion of the Simulated Aggregated Loss with Actual Loss   

```{r}
library(gridExtra)
Aggregated_Loss=c(summary(total_cost),skewness=round(skewness(total_cost),1),
                  kurtosis=round(kurtosis(total_cost),1))
Acutal_Loss=c(summary(Total$Total),skewness=round(skewness(Total$Total),1),
                  kurtosis=round(kurtosis(Total$Total),1))
g=data.frame(Aggregated_Loss,Acutal_Loss)
g=tableGrob(g)
grid.draw(g)
```

###Actual Aggregated Loss
```{r}
hist(Total$Total,breaks = 20,freq = F,
     main="Actual Aggregated Loss",
     col="orchid",xaxt="n",#yaxt="n",
     xlab="Total Cost (in millions)")

xrange=seq.int(0,3.5e7,by = 5e6)
xlabel=paste("$",xrange/1e6,sep="")
axis(1,at=xrange,labels = xlabel)
```        
      
###Diagnostic of the Aggregated Loss Model Against Actual Aggregated Loss    

```{r}
rand_lnorm=rlnorm(1e5,meanlog = lnorm_param$estimate["meanlog"],
                  sdlog = lnorm_param$estimate["sdlog"])
exprange=c(0,2e7)
qqplot(Total$Total,rand_lnorm,col="green",pch=19,xlim=exprange,ylim=exprange,
       ylab="Aggregated Loss Model",xlab="Actual Aggregated Loss",main="QQ-plot")
abline(0,1,col="red",lty=3,lwd=3)
```

###Kolmogorov-Smirnov (K-S) test     
```{r}
ks.gof(Total$Total,"plnorm",meanlog = lnorm_param$estimate["meanlog"],
       sdlog = lnorm_param$estimate["sdlog"])
```

###VaR: 

**The Value at Risk at 95% confident Interval for one year aggregated loss is `r paste0("$",round(quantile(total_cost,c(.95))/1e6))` millions of dollars**    

**The Value at Risk at 99% confident Interval for one year aggregated loss is `r paste0("$",round(quantile(total_cost,c(.99))/1e6))` millions of dollars**   

###Expected Shortfall   
**The expected amount of aggregated loss given that the loss is beyond the confident Interval**   
```{r}
slice = 5e4
slice.seq=seq(from = .95,to = 1,length.out = slice)

```

**The Expected Shortfall at 95% confident Interval is `r paste0("$",round(mean(quantile(total_cost,slice.seq))/1e6))` millions of dollars** 

```{r}
slice.seq=seq(from = .99,to = 1,length.out = slice)
```

**The Expected Shortfall at 99% confident Interval is `r paste0("$",round(mean(quantile(total_cost,slice.seq))/1e6))` millions of dollars**  


[EEPC]:http://www.nyc.gov/html/eepc/html/home/annual-reports.shtml